---
title: "Code_tvQAR"
author: "Manon Felix"
date: "2023-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [1.] Packages

```{r}
library(roahd)
library(fda)
library(rootSolve)
library(ggplot2)
library(hrbrthemes)
library(tidyverse)
library(gridExtra)
library(grid)
library(gridtext)
```


# [2.] Functions 


## [2.1] Kernel function and check function

```{r}
# Define a function called "kern.func" which takes in three arguments: "u0", "n", and "b".
kern.func <- function(u0, n, b) {

u <- (u0 - (1:n)/n)/b

# Calculate the value of "K" using the ifelse statement. 
# -If the absolute value of "u" is less than or equal to 1, K will equal 0.75 times (1 minus the absolute value of u squared). 
# -If the absolute value of "u" is greater than 1, K will equal 0.
K = ifelse(abs(u) <= 1, 0.75*(1 - abs(u)^2), 0)

# Return the value of "K".
return(K)
}

rho <- function(u, tau){u*(tau - as.numeric(u <= 0))}
```


## [2.2] Loss functions 

* 1. Loss without the kernel 
* 2. Loss with constant polynomial approximation
* 3. Loss with linear polynomial approximation
* 4. Loss with quadratic polynomial approximation

```{r}
loss_nocst <- function(u0, tau, theta = c(), T, b, p){
  
  Y_vec = na.omit(tsutils::lagmatrix(Y, -p)) #matrix of size T-p x p
  
  lagmat = na.omit(tsutils::lagmatrix(Y, 1:p))
  
  covariates_mat = cbind(rep(1, T-p), lagmat) #matrix of size T-p x (p + 1) because we add the interecept column 
  
  u = Y_vec - as.numeric(covariates_mat %*% theta[1:(p+1)]) #If p = 3, T = 10, Y starts from Y4 to Y10 

  loss = sum(rho(u = u, tau = tau)) #multiply the loss by the kernel 
  
  return(loss)
}

loss_const <- function(u0, tau, theta = c(), T, b, p){
  
  Y_vec = na.omit(tsutils::lagmatrix(Y, -p)) #matrix of size T-p x p
  
  lagmat = na.omit(tsutils::lagmatrix(Y, 1:p))
  
  covariates_mat = cbind(rep(1, T-p), lagmat) #matrix of size T-p x (p + 1) because we add the interecept column 
  
  u = Y_vec - as.numeric(covariates_mat %*% theta[1:(p+1)]) #If p = 3, T = 10, Y starts from Y4 to Y10 
  
  Ki_u0 = kern.func(u0, (T-p), b) # vector (n - p) x 1 
  
  loss = sum(rho(u = u, tau = tau)*kern.func(u0, (T-p), b)) #multiply the loss by the kernel 
  
  return(loss)
}

loss_linear <- function(u0, tau, theta = c(), T, b, p){
  
  Y_vec = na.omit(tsutils::lagmatrix(Y, -p)) #matrix of size T-p x p
  
  lagmat = na.omit(tsutils::lagmatrix(Y, 1:p))
  
  covariates_mat = cbind(rep(1, T-p), lagmat) #matrix of size T-p x (p + 1) because we add the interecept column 
  
  i = (1+p):(T)
  
  val = i/T - u0 # terms which multiply the local linear part 
  
  lin_part_coef = theta[(p+2):(2*(p+1))] 
  
  dim(lin_part_coef) = c(p+1,1) 
  
  dim(theta) = c(2*(p+1), 1)
  
  us = Y_vec - as.numeric(covariates_mat %*% theta[1:(p+1)]) - as.numeric(covariates_mat %*% lin_part_coef * val)
  
  L = sum(rho(us, tau) * kern.func(u0, (T-p), b))
  
  return(L)
}


loss_quadra <- function(u0, tau, theta = c(), T, b, p){
  
  Y_vec = na.omit(tsutils::lagmatrix(Y, -p)) #matrix of size T-p x p
  
  lagmat = na.omit(tsutils::lagmatrix(Y, 1:p))
  
  covariates_mat = cbind(rep(1, T-p), lagmat) #matrix of size T-p x (p + 1) because we add the interecept column 
  
  i = (1+p):(T)
  
  val = i/T - u0 # terms which multiply the local linear part 
  
  lin_part_coef = theta[(p+2):(2*(p+1))] 
  
  quadra_part_coef = theta[(p + 1 + p + 1 + 1):(3*(p+1))] 
  
  dim(lin_part_coef) = c(p+1,1) 
  
  dim(quadra_part_coef) = c(p+1,1) 
  
  dim(theta) = c(3*(p+1), 1)
  
  us = Y_vec - as.numeric(covariates_mat %*% theta[1:(p+1)]) - as.numeric(covariates_mat %*% lin_part_coef * val) -  as.numeric(covariates_mat %*% quadra_part_coef * val^2 / 2)
  
  L = sum(rho(us, tau) * kern.func(u0, (T-p), b))
  
  return(L)
}
```


## [2.3] Parameters function

```{r}
theta_fct_sine <- function(t, T, c = 18, d= 0.2) {
    (d*sin(t/T * c) + 0.2 +  c/30 * (t / T) + d/25 * (t / T) - 0.08*((t/T)+1)^3)*0.4
}
```


# [3.] Code Figure 1. 

```{r}
## Parabolic 
theta_fct_para <- function(t, T, c = 0.1, d = 0.85) {
 c * (t / T) + d * (t / T)^{2.5}
 # c * (t / T) + d * (t / T)^3
}


plot(1:100/100, theta_fct_para(1:100,100), type = "l")
```


```{r, eval = FALSE}
######################### 0.5 #######################


m = mt = 1
n = 4000
dim_para = p = 1
u0s = seq(0,1,l=(n-1))
set.seed(29)

Q_nocst5 <- matrix(NA, nrow = mt, ncol = (length(u0s)))
par_nocst <- array(rep(1, mt*(length(u0s))*(dim_para+1)), dim=c(mt, length(u0s), dim_para+1))




 # Simulate the AR(1)
  eps = rnorm(n)
  
  Y <- array(NA, dim = n)
  Y[1:dim_para] <- 0 
  
  for (i in dim_para:(n-1)){
  
  Y[i + 1] = rep(as.numeric(0.5, dim_para)) %*% Y[(i-(dim_para-1)):i] + eps[i]

  }

  # Use the mapply function instead of a for loop to optimize coef and coef_lin
  results <- mapply(function(j) {
    list(
      #optim(par = c(rep(0, dim_para+1)), loss_const, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par,
      optim(par = c(rep(0, dim_para+1)), loss_nocst, tau = 0.5, u0 = u0s[j], T = n, b = b, p = dim_para)$par
    )
  }, j = 1:(length(u0s)))
  
  par_nocst[m,,] = do.call(rbind, results)

  ## Scatterplot; estimated quantiles / true quantiles
  lagmat = t(cbind(rep(1, n-p), na.omit(tsutils::lagmatrix(Y, 1:p))))
  #Q0[m,] = c(qnorm(taus,0,1) + theta_fct_para(u0s[j]*n, T = n) * Y[val_Y])
  Q_nocst5[m,] = as.numeric(diag(par_nocst[m,,] %*% lagmat))
  #Q_lin[m,] = diag(par_lin[m,,1:(p+1)] %*% lagmat)



theta_true = cbind(rep(qnorm(0.5,0,1), n-1), 0.5)

Q5_truecst = diag(theta_true %*% lagmat)

Q_nocst05 <- matrix(NA,  nrow = mt, ncol = (length(u0s)))

  # Simulate the AR(1)
  
  Y <- array(NA, dim = n)
  Y[1:dim_para] <- 0 
  
  for (i in dim_para:(n-1)){
  
  Y[i + 1] = rep(as.numeric(theta_fct_para(i+1, n)), dim_para) %*% Y[(i-(dim_para-1)):i] + eps[i]

  }
  # Use the mapply function instead of a for loop to optimize coef and coef_lin
  results <- mapply(function(j) {
    list(
      optim(par = c(rep(0, dim_para+1)), loss_nocst, tau = 0.5, u0 = u0s[j], T = n, b = b, p = dim_para)$par
    )
  }, j = 1:(length(u0s)))
  
  par_nocst[m,,] = do.call(rbind, results)

  ## Scatterplot; estimated quantiles / true quantiles
  lagmat = t(cbind(rep(1, n-p), na.omit(tsutils::lagmatrix(Y, 1:p))))
  #Q0[m,] = c(qnorm(taus,0,1) + theta_fct_para(u0s[j]*n, T = n) * Y[val_Y])
  Q_nocst05[m,] = as.numeric(diag(par_nocst[m,,] %*% lagmat))


theta_true = cbind(rep(qnorm(0.5,0,1), n-1), theta_fct_para(u0s*n, n))

Q05_truetv = diag(theta_true %*% lagmat)

# A basic scatterplot with color depending on Species
p5 <- ggplot(NULL, aes(x=Q5_truecst, y=Q_nocst5)) + 
  geom_point(aes(x=Q05_truetv, y=Q_nocst05), color="#7d9fc2", shape = 3, size = 1) +
  geom_abline(slope = 1, size = 1.2) +
  geom_point(size=1, color = "#C582B2", shape = 4)+
  theme_minimal()+
  labs(x = "True conditional quantiles", y = "Estimated conditional quantiles") 
  #coord_fixed(xlim = c(min(Q_nocst05), max(Q_nocst05)))


######################### 0.15 #######################


dim_para = p = 1
u0s = seq(0,1,l=(n-1))
set.seed(29)

Q_nocst15 <- matrix(NA,  nrow = mt, ncol = (length(u0s)))
par_nocst <- array(rep(1, mt*(length(u0s))*(dim_para+1)), dim=c(mt, length(u0s), dim_para+1))


m = 1
  
  Y <- array(NA, dim = n)
  Y[1:dim_para] <- 0 
  
  for (i in dim_para:(n-1)){
  
  Y[i + 1] = rep(as.numeric(0.5, dim_para)) %*% Y[(i-(dim_para-1)):i] + eps[i]

  }

  # Use the mapply function instead of a for loop to optimize coef and coef_lin
  results <- mapply(function(j) {
    list(
      #optim(par = c(rep(0, dim_para+1)), loss_const, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par,
      optim(par = c(rep(0, dim_para+1)), loss_nocst, tau = 0.15, u0 = u0s[j], T = n, b = b, p = dim_para)$par
    )
  }, j = 1:(length(u0s)))
  
  par_nocst[m,,] = do.call(rbind, results)

  ## Scatterplot; estimated quantiles / true quantiles
  lagmat = t(cbind(rep(1, n-p), na.omit(tsutils::lagmatrix(Y, 1:p))))
  #Q0[m,] = c(qnorm(taus,0,1) + theta_fct_para(u0s[j]*n, T = n) * Y[val_Y])
  Q_nocst15[m,] = as.numeric(diag(par_nocst[m,,] %*% lagmat))
  #Q_lin[m,] = diag(par_lin[m,,1:(p+1)] %*% lagmat)



theta_true = cbind(rep(qnorm(0.15,0,1), n-1), 0.5)

Q15_truecst = diag(theta_true %*% lagmat)

Q_nocst015 <- matrix(NA,  nrow = mt, ncol = (length(u0s)))

  # Simulate the AR(1)
  
  Y <- array(NA, dim = n)
  Y[1:dim_para] <- 0 
  
  for (i in dim_para:(n-1)){
  
  Y[i + 1] = rep(as.numeric(theta_fct_para(i+1, n)), dim_para) %*% Y[(i-(dim_para-1)):i] + eps[i]

  }
  # Use the mapply function instead of a for loop to optimize coef and coef_lin
  results <- mapply(function(j) {
    list(
      optim(par = c(rep(0, dim_para+1)), loss_nocst, tau = 0.15, u0 = u0s[j], T = n, b = b, p = dim_para)$par
    )
  }, j = 1:(length(u0s)))
  
  par_nocst[m,,] = do.call(rbind, results)

  ## Scatterplot; estimated quantiles / true quantiles
  lagmat = t(cbind(rep(1, n-p), na.omit(tsutils::lagmatrix(Y, 1:p))))
  #Q0[m,] = c(qnorm(taus,0,1) + theta_fct_para(u0s[j]*n, T = n) * Y[val_Y])
  Q_nocst015[m,] = as.numeric(diag(par_nocst[m,,] %*% lagmat))


theta_true = cbind(rep(qnorm(0.15,0,1), n-1), theta_fct_para(u0s*n, n))

Q015_truetv = diag(theta_true %*% lagmat)

# A basic scatterplot with color depending on Species
p15 <- ggplot(NULL, aes(x=Q15_truecst, y=Q_nocst15)) + 
  geom_point(aes(x=Q015_truetv, y=Q_nocst015), color="#7d9fc2", shape = 3, size = 1) +
  geom_abline(slope = 1, size = 1.2) +
  geom_point(size=1, color = "#C582B2", shape = 4)+
  theme_minimal()+
  labs(x = "True conditional quantiles", y = "Estimated conditional quantiles") 
  #coord_fixed(xlim = c(min(Q_nocst05), max(Q_nocst05)))



# Remove axis titles from all plots
p = list(p15, p5) %>% map(~.x + labs(x=NULL, y=NULL))
# gridtext
yleft = richtext_grob("Estimated conditional quantiles", rot=90)

bottom = richtext_grob(
  text = 'True conditional quantiles'
)

# Lay out plots
uni <- grid.arrange(grobs=p, ncol = 2, nrow = 1, left = yleft, bottom = bottom)
```

# [4.] Monte-Carlo simulations 

```{r, eval = FALSE}
set.seed(1)
mt = 100

dim_para = p = 3

tau = 0.5

n = 3000
b = 0.05

u0s = seq(0, 1, l = n/30)
j = 1:length(u0s)

par_cst <- array(rep(1, mt*length(u0s)*(dim_para+1)), dim=c(mt, length(u0s), dim_para+1))
par_lin <- array(rep(1, mt*length(u0s)*(2*dim_para+2) ),dim=c(mt, length(u0s), 2*dim_para+2))
par_quadra <- array(rep(1, mt*length(u0s)*(3*(dim_para+1)) ),dim=c(mt, length(u0s), 3*(dim_para+1)))


Q0 <- matrix(NA,  ncol = mt, nrow = length(u0s))
Q_cst <- matrix(NA,  ncol = mt, nrow = length(u0s))
Q_lin <- matrix(NA,  ncol = mt, nrow = length(u0s))


for (m in 1:mt){
  
  # Simulate the AR(10)
  eps = rnorm(n)
  
  Y <- array(NA, dim = n)
  Y[1:dim_para] <- 0 
  
  for (i in dim_para:(n-1)){
  
   Y[i + 1] = as.numeric(c(theta_fct_sine(-(i+1)+1, n)/3, theta_fct_sine(-(i+1)+1, n), theta_fct_sine(-(i+1)+1, n)/10)) %*% Y[(i-(dim_para-1)):i] + eps[i]

  }
  
  # Use the mapply function instead of a for loop to optimize coef and coef_lin
  results <- mapply(function(j) {
    list(
      optim(par = c(rep(0, dim_para+1)), loss_const, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par,
      optim(par = c(rep(0, 2*(dim_para+1))), loss_linear, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par,
      optim(par = c(rep(0, 3*(dim_para+1))), loss_quadra, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par
    )
  }, j = 1:(length(u0s)))
  
  par_cst[m,,] = do.call(rbind, results[seq(1,(3*length(u0s)-2), by = 3)])
  par_lin[m,,] = do.call(rbind, results[seq(2,(3*length(u0s)-1), by = 3)])
  par_quadra[m,,] = do.call(rbind, results[seq(3,(3*length(u0s)), by = 3)])

  print(m)
}

# save(par_cst, file="par_cst.rda")
# save(par_lin, file="par_lin.rda")
# save(par_quadra, file="par_quadra.rda")
```

## [4.1] Code Figure 2. 

Load the results from the Monte-Carlo simulations. 

```{r}
load("par_cst.rda")
load("par_lin.rda")
load("par_quadra.rda")
```

Note that we changed the colors and y-scale by modifiying directly the source's code. 

```{r}
n = 3000
u0s = seq(0, 1, l = n/30)
##### Functional boxplots
op <- par(mfrow = c(3,1),
          oma = c(4,3,0,0) + 0.1,
          mar = c(0,0,2,0) + 0.1)

theta2 = theta_fct_sine(-u0s*n + 1,n)

grid2 = u0s
fD = fData(grid2, par_cst[,,3])

roahd::fbplot(fD, 
        xlab = '', ylab = '', 
           Fvalue = 10, lwd =1,
        main = '')
lines(u0s, theta2)

grid2 = u0s
fD = fData(grid2, par_lin[,,3])

roahd::fbplot(fD, 
        xlab = '', ylab = '', 
           Fvalue = 10, lwd =1,
        main = '')
lines(u0s, theta2)

fD = fData(grid2, par_quadra[,,3])

roahd::fbplot(fD, 
        xlab = 'i/n', ylab = '', 
           Fvalue = 10, lwd =1,
        main = '')
lines(u0s, theta2)
```

## [4.2] Code Figure 3. 

```{r}
MSE_lin <- mapply(function(j){
  mean((par_lin[,j,3] - theta_fct_sine(-u0s[j]*n+1,n))^2)
}, j = 1:(length(u0s)))

BIAS_lin = mapply(function(j){
  mean((par_lin[,j,3] - theta_fct_sine(-u0s[j]*n+1,n)))
}, j = 1:(length(u0s)))

MSE_cst <- mapply(function(j){
  mean((par_cst[,j,3] - theta_fct_sine(-u0s[j]*n+1,n))^2)
}, j = 1:(length(u0s)))

BIAS_cst = mapply(function(j){
  mean((par_cst[,j,3] - theta_fct_sine(-u0s[j]*n+1,n)))
}, j = 1:(length(u0s)))

MSE_quadra <- mapply(function(j){
  mean((par_quadra[,j,3] - theta_fct_sine(-u0s[j]*n+1,n))^2)
}, j = 1:(length(u0s)))

BIAS_quadra = mapply(function(j){
  mean((par_quadra[,j,3] - theta_fct_sine(-u0s[j]*n+1,n)))
}, j = 1:(length(u0s)))


a = data.frame(group = "Local constant", value = MSE_cst)
b = data.frame(group = "Local linear", value =  MSE_lin)
c = data.frame(group = "Local quadratic", value = MSE_quadra)

plot.data = rbind(a, b, c) 

plot.data$u0s <- rep(u0s, 3)


ggplot(plot.data, aes(x = u0s, y = value, group = group, col = group, linetype = group)) +
  geom_line(lwd = 1) +
  scale_linetype_manual(values = c("dashed", "solid", "dotted")) +
  scale_color_manual(values = c("#7d9fc2","#DEB478", "#C582B2" )) +
  theme_bw() +
  labs(x = "i/n", y = "Estimated MSE") +
  theme(legend.title= element_blank())

```

# [5.] Computation time 

```{r}
set.seed(1)
library(tictoc)
mt = 100

dim_para = p = 3


## Sine wave 
theta_fct_sine <- function(t, T, c = 38, d= 0.2) {
  (d*sin(t/T * c) + 0.2  * (t / T) + d/25 * (t / T) - 0.08*((t/T)+1)^3) + 0.5
  
}

tau = 0.5

n = 3000
b = 0.05

u0s = seq(0, 1, l = n/30)
j = 1:length(u0s)


  eps = rnorm(n)
  
  Y <- array(NA, dim = n)
  Y[1:dim_para] <- 0 
  
  for (i in dim_para:(n-1)){
    
    Y[i + 1] = as.numeric(c(theta_fct_sine(-(i+1)+1, n)/3, theta_fct_sine(-(i+1)+1, n), theta_fct_sine(-(i+1)+1, n)/10)) %*% Y[(i-(dim_para-1)):i] + eps[i]
    
  }
  

  # Estimation 
tic()
      mapply(function(j){
      optim(par = c(rep(0, dim_para+1)), loss_const, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par
      }, j = 1:(length(u0s)))
toc()
      
tic()
      mapply(function(j){
      optim(par = c(rep(0, 2*(dim_para+1))), loss_linear, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par
      }, j = 1:(length(u0s)))
toc()  

tic()
mapply(function(j){
       optim(par = c(rep(0, 3*(dim_para+1))), loss_quadra, tau = tau, u0 = u0s[j], T = n, b = b, p = dim_para)$par
}, j = 1:(length(u0s)))
toc()  
  
```

